Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-09-01 main.tex
review typos

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)
\noindent
Let $\Semiring$ be a commutative semiring,
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets},
and $\bar\Phi$ be a label theory over $\Semiring$
containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,
and $\bar{\wei}$ is a triplet of transition functions
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call \emph{number of transitions} of $T$ the number of pairs of states
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions
as functions of
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for all $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) &
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}
\]
%
\noindent
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$),
or the opposite ($\wei_{01}$),
or advance in both $s$ and $t$ ($\wei_{11}$).}
% \noindent
$T$ is
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$,
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
% \setminus \{ \< \varepsilon, \varepsilon >\}$,
%
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
 \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}}
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}}
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}}
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}

%
We recall that, by convention (Section~\ref{sec:prelim}),
an empty sum with $\bigoplus$ is equal to~$\zero$.
%
Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
when reading $a$ and $b$ in the input and output words,
increment the current position in the input word if and only if $i = 1$,
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions,
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
symbolic transducers.

@@

The expression \eqref{eq:SWT-weight}
can be seen as a stateful definition of
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively,
$\wei_{10}(q, a, \varepsilon, r)$ is the cost of
the deletion of the symbol~$a \in \Sigma$ in~$s$,
$\wei_{01}(q, \varepsilon, b, r)$ is the cost
of the insertion of~$b \in \Delta$ in $t$,
and $\wei_{11}(q, a, b, r)$ is the cost
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$,
is the product, with $\otimes$, of the individual costs of the operations involved;
and the distance between $s$ and $t$ is the sum, with $\oplus$,
of all possible products.
%
\medskip\noindent
%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$.
Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is:
%defined as follows:
\begin{equation}
T(s, t)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
\label{eq:SWT-value}
\end{equation}

@@

\begin{example}\label{ex:SWT}
We build a small \SWT over  alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{}),
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
(respectively $\mathsf{out} : Q \to \Semiring$)  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based with an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q',\ 
          \mathrm{or}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
there exists a \SWA $B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
effectively constructible in PTIME, 
such that for all $t \in \Delta^+$,
$B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
\end{proposition}
%
\begin{proof}
(sketch, see Appendix~\ref{sec:closure} for details).
The state set of $B_{A, T}$ is the Cartesian product of
the state sets of $A$ and $T$ and its transitions simulate,
while reading an output word $t \in \Delta^*$, the
synchronized behaviour of $A$ and $T$ on $t$
and some input word $s \in \Sigma^*$.
The weight for reading the input $s$ is obtained with
$\bigoplus^1_\Sigma$.
%
The main difficulty comes from the transitions
of the form $\wei_{10}$, which read in input and ignore the output.
Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
but $\varepsilon$-transitions are not defined for $\SWA$.
Therefore, the $\varepsilon$-transitions
are eliminated on-the-fly during the construction of~$B_{A, T}$,
following a procedure of~\cite{LombardySakarovitch12ciaa}.
%\qed
\end{proof}

%The complexity of construction
%\noindent
%The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
%where the sizes $\| T \|$ and $\| A \|$ are their number of states.
%%of~$T$ is its number of states $|Q|$.
%\florent{revise complexity with nb. of tr. and states? 
%         or skip details and just say polynomial?}
%
The particular case of Proposition~\ref{prop:epsilon} with a singleton $A$,
\ie such that $A(s) = \one$ for a given $s \in \Sigma^*$
and $A(s') = \zero$ for all $s' \neq s$,
corresponds to a construction of a \SWA for the partial application of the \SWT $T$,
fixing the first argument $s$.
\begin{corollary} \label{cor:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and $s \in \Sigma^+$,
there exists an effectively constructible \SWA
$B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
\end{corollary}

@@

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Delta$ be a countable alphabet
%finite (large) or infinite,
artitioned into three
subsets~$\Deltai$, $\Deltac$, $\Deltar$,
whose elements are respectively called
\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
%we shall write respectively
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
stand respectively
for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$
%of unary functions in $\Deltai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.

>>>>

%\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)
\noindent
Let $\Semiring$ be a commutative semiring,
$\Sigma$ and $\Delta$ be alphabets called \emph{input} and \emph{output} respectively, %{alphabets},
and $\bar\Phi$ be a label theory over $\Semiring$
containing $\Phi_\Sigma$, $\Phi_\Delta$, $\Phi_{\Sigma, \Delta}$.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} \colon Q \to \Semiring$   %\Phi_{\Sigma, \Delta}
(respectively $\mathsf{out} \colon Q \to \Semiring$)  %\Phi_{\Sigma, \Delta}
are functions defining the weight for entering
(respectively leaving) computation in a state,
and $\bar{\wei}$ is a triplet of transition functions
$\wei_{10}: Q \times Q \to \Phi_{\Sigma}$,
$\wei_{01}: Q \times Q \to \Phi_{\Delta}$, and
$\wei_{11}: Q \times Q \to \Phi_{\Sigma, \Delta}$.
\end{definition}
%
\noindent
We call the \emph{number of transitions} of $T$ the number of pairs of states
$q, q' \in Q$ such that $\wei_{10}$ or $\wei_{01}$ or $\wei_{11}$
is not the constant $\zero$.
%
\noindent
For convenience, we shall sometimes present transitions
as functions of
$Q \times (\Sigma \cup \{ \varepsilon \}) \times (\Delta \cup \{ \varepsilon \}) \times Q \to \Semiring$,
overloading the function names,
such that, for every $q, q' \in Q$, $a \in \Sigma$,  $b \in \Delta$:
\[
\begin{array}{rcll}
\wei_{10}(q, a, \varepsilon, q') & = & \phi(a) & %\wei_\Sigma(q, q')(a)
\quad\mathrm{where~} \phi = \wei_{10}(q, q') \in \Phi_\Sigma,\\
\wei_{01}(q, \varepsilon, b, q') & = & \psi(b) &
\quad\mathrm{where~} \psi = \wei_{01}(q, q') \in \Phi_\Delta,\\
\wei_{11}(q, a, b, q') & = & \eta(a, b) &
\quad\mathrm{where~} \eta = \wei_{11}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}
\]
%
\noindent
% The symbolic-weighted transducer
The \SWT $T$ computes on pairs $\< s, t> \in \Sigma^* \times \Delta^*$,
$s$ and $t$, being respectively called \emph{input} and \emph{output} word.
\florent{It compute asynchronously: advance in $s$ and not in $t$ ($\wei_{10}$),
or the opposite ($\wei_{01}$),
or advance in both $s$ and $t$ ($\wei_{11}$).}
% \noindent
$T$ is
based on an intermediate function $\weight_T$
defined recursively, for every states $q, q' \in Q$,
and every pairs of strings $\< s, t> \in \Sigma^* \times \Delta^*$.
% \setminus \{ \< \varepsilon, \varepsilon >\}$,
%
%$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, by:
%
\begin{align}
%\weight_T(q, \varepsilon, \varepsilon, q)  & = \one \label{eq:SWT-weight}\\ %\final(q)\\
%\weight_T(q, \varepsilon, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_T(q, \varepsilon, \varepsilon, q')  & = \one
 \quad \mathrm{if~} q = q' \mathrm{~and~} \zero \mathrm{~otherwise} \label{eq:SWT-weight}\\
\weight_T(q, s, t, q') & = \displaystyle\bigoplus_{\begin{array}{c}
                                                   \scriptstyle q'' \in Q\\[-1ex]
                                                   \scriptstyle s = au,\, a \in \Sigma
                                                   \end{array}}
    \wei_{10}(q, a, \varepsilon, q'') \otimes \weight_T(q'', u, t, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle t = bv,\, b \in \Delta\\
                                                     \end{array}}
    \wei_{01}(q, \varepsilon, b, q'') \otimes \weight_T(q'', s, v, q')\nonumber\\
                    & \oplus \displaystyle\bigoplus_{\begin{array}{c}
                                                     \scriptstyle q'' \in Q\\[-1ex]
                                                     \scriptstyle s = au,\, t = bv\\
                                                     \end{array}}
    \wei_{11}(q, a, b, q'') \otimes \weight_T(q'', u, v, q')\nonumber
\end{align}

%
We recall that, by definition (Section~\ref{sec:prelim}),
an empty sum with $\bigoplus$ is equal to~$\zero$.
%
Intuitively,  a transition $\wei_{ij}(q, a, b, q')$ is interpreted as follows:
when reading $a$ and $b$ in the input and output words,
increment the current position in the input word if and only if $i = 1$,
and in the output word iff $j = 1$, %(otherwise, do not change it),
and change state from $q$ to $q'$.
When $a = \varepsilon$ (resp. $b = \varepsilon$), the current symbol
in the input (resp. output) is not read.
%
%In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%the input and output symbols at current positions are always read by transitions,
%even when they do not change the reading position the head's position.
%This is an important feature in the case of an infinite alphabet in
%order to compare input and output symbols.
%which cannot be stored in the finite memory of the transducer.
%
Since~$\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for
symbolic transducers.

@@

The expression \eqref{eq:SWT-weight}
can be seen as a stateful definition of
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively,
$\wei_{10}(q, a, \varepsilon, r)$ is the cost of
the deletion of the symbol~$a \in \Sigma$ in~$s$,
$\wei_{01}(q, \varepsilon, b, r)$ is the cost
of the insertion of~$b \in \Delta$ in $t$,
and $\wei_{11}(q, a, b, r)$ is the cost
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$
is the product in terms of $\otimes$ of the individual costs of the operations involved;
and the distance between $s$ and $t$ is the sum in terms of $\oplus$
of all possible products.
%
\medskip\noindent
%Let $\< s, t> \in \Sigma^* \times \Delta^*$, with $s = s_1\ldots s_n$, and $t = t_1\ldots t_m$.
Formally, the weight associated by $T$ to $\< s, t> \in \Sigma^* \times \Delta^*$ is:
%defined as follows:
\begin{equation}
T(s, t)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \final(q')
\label{eq:SWT-value}
\end{equation}

@@

\begin{example}\label{ex:SWT}
We build a small \SWT over the alphabets $\Sigma$ and $\Deltai$
(Ex.~\ref{ex:running} and \ref{distance-time}),
%weighted transducer model
with two states $q_0$ and $q_1$, that
calculates the temporal distance between an input performance in~$\Sigma^*$
and the subsequence
of $\Deltai$ events in a score.
%the performance $I$ and
%the score $O$ of Example~\ref{ex:running}.
%
Given  a performed event $\mu$
and the corresponding notated event $\nu$
(\eg MIDI pitch 69 and note $\mbox{\footnotesize A4}$),
the weight  computed by the \SWT is the time distance between both,
as modeled by transitions $\wei_{11}$ below.
%
The continuation symbol $'-'$
(met for instance in \emph{ties}
$\musQuarter\!\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\!\musEighth$,
or \emph{dots} \musQuarterDotted{})
%that represents continuation
is  skipped with no cost (transitions $\wei_{01}$).
%Likewise, the markups are ignored.
\[
\begin{array}{lclclcll}
\wei_{11}(q_0, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) & \quad & %|\tau' - \tau| & \quad &
\wei_{11}(q_1, \mu\ts{\tau}, \nu\ts{\tau'}, q_0) & = &
     \delta(\mu\ts{\tau}, \nu\ts{\tau'}) &
     \mbox{if~} \nu \neq - \\ %|\tau' - \tau| & \quad &
\wei_{01}(q_0, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one & &
\wei_{01}(q_1, \varepsilon, \mathsf{-}\ts{\tau'}, q_0) & = & \one\\
\wei_{10}(q_0, \mu\ts{\tau}, \varepsilon, q_1) & = & \alpha & & %\multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
We also want to take performing errors into account,
since a performer could, for example, play an unwritten extra note.
%\lydia{reformulated this sentence}
%
%Reading such a note is modelled by
The transition $\wei_{10}$,
with an fixed weight value $\alpha \in \Semiring$,
switches from state $q_0$ (normal) to $q_1$ (error)
when reading an extra note $\mu$.
%\philippe{Comprends pas cette phrase}
%\florent{$\to$ reformulation}
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \final(q_0) = \one$, and
$\init(q_1) = \final(q_1) = \zero$.
%\florent{dernière phrase utile?}
%This \SWT is capable of evaluating the differences between a score and a performance,
%all the while ensuring that performance errors are plausible.
%$\init(q_0, d, b) = \final(q_0, d, b) = \one$, and
%$\init(q_1, d, b) = \final(q_1, d, b) = \zero$,
%for all $d \in \Sigma$ and $b \in \Delta$.
\endex
\end{example}

@@

\noindent
 \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined  as the transducers of Definition~\ref{def:SWT},
by simply omitting the output symbols.
%
%In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$
is a tuple
$A = \< Q, \init, {\wei_1}, \final >$,
where $Q$ is a finite set of states,
$\mathsf{in} : Q \to \Semiring$ %\Phi_\Sigma$,
, respectively $\mathsf{out} : Q \to \Semiring$,  %\Phi_\Sigma,$
are functions defining the weight for entering
(respectively leaving) computation in a state,
and ${\wei_1}$ is a transition function
from $Q \times Q$ into~$\Phi_{\Sigma}$.
\end{definition}
%
\noindent
%\florent{mv formal def. weight for \SWA to app.?}
As above in the case of \SWT,
when $\wei_1(q, q') = \phi \in \Phi_\Sigma$,
%respectively $\mathsf{in}(q) = \phi$, $\mathsf{out}(q') = \phi$,
we may write $\wei_1(q, a, q')$ for~$\phi(a)$.
%respectively $\mathsf{in}(q, a) = \phi(a)$, $\mathsf{out}(q', a) = \phi(a)$.
%$\wei_1: Q \times \Sigma \times Q \to \Semiring$,
The computation of $A$ on words $s \in \Sigma^*$
is based on with an intermediate function $\weight_A$,
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$
%
\begin{align}
\weight_A(q, \varepsilon, q') & = \one\ \mathrm{if}\  q = q'\ 
          \mathrm{and}\  \zero\  \mathrm{otherwise}\\
%\weight_A(q, \varepsilon, q') & = \zero \quad \mathrm{if~} q \neq q'\nonumber\\
\weight_A(q, au, q') & =  \displaystyle\bigoplus_{q'' \in Q}
    \wei_{1}(q, a, q'') \otimes \weight_A(q'', u, q')\nonumber
\label{eq:SWA-weight}
\end{align}
%
\noindent
and the weight value associated by $A$ to
$s \in \Sigma^*$ is defined as follows: %$s = s_1\ldots s_n \in \Sigma^+$
\begin{equation}
A(s)  =
\displaystyle\bigoplus_{q, q' \in Q} \init(q)
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \final(q')
\label{eq:SWA-value}
\end{equation}

@@

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and a $\SWA$ $A$ over $\Sigma$, $\Semiring$ and $\bar\Phi$,
there exists a \SWA $B_{A, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
effectively constructible in PTIME, 
such that for every $t \in \Delta^+$,
$B_{A, T}(t) = \displaystyle\bigoplus_{s\in \Sigma^*} A(s) \otimes T(s, t)$.
\end{proposition}
%
\begin{proof}
(sketch, see Appendix~\ref{sec:closure} for details).
The state set of $B_{A, T}$ is the Cartesian product of
the state sets of $A$ and $T$ and its transitions simulate,
while reading an output word $t \in \Delta^*$, the
synchronized behaviour of $A$ and $T$ on $t$
and some input word $s \in \Sigma^*$.
The weight for reading the input $s$ is obtained with
$\bigoplus^1_\Sigma$.
%
The main difficulty comes from the transitions
of the form $\wei_{10}$, which read in input and ignore the output.
Such transition shall be simulated by $\varepsilon$-transitions in $B_{A, T}$,
but $\varepsilon$-transitions are not defined for $\SWA$.
Therefore, the $\varepsilon$-transitions
are eliminated on-the-fly during the construction of~$B_{A, T}$,
following a procedure of~\cite{LombardySakarovitch12ciaa}.
%\qed
\end{proof}

%The complexity of construction
%\noindent
%The construction time and size for $B_{A, T}$ are $O(\| T \|^3 . \| A \|^2)$,
%where the sizes $\| T \|$ and $\| A \|$ are their number of states.
%%of~$T$ is its number of states $|Q|$.
%\florent{revise complexity with nb. of tr. and states? 
%         or skip details and just say polynomial?}
%
The particular case of Proposition~\ref{prop:epsilon} with a singleton $A$,
\ie such that $A(s) = \one$ for a given $s \in \Sigma^*$
and $A(s') = \zero$ for every $s' \neq s$,
corresponds to a construction of a \SWA for the partial application of the \SWT $T$,
fixing the first argument $s$.
\begin{corollary} \label{cor:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$,
$\Semiring$ commutative, bounded and complete,
and $\bar\Phi$ effective,
and $s \in \Sigma^+$,
there exists an effectively constructible \SWA
$B_{s, T}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for every $t \in \Delta^+$, $B_{s, T}(t) = T(s, t)$.
\end{corollary}

@@

%\subsection{Definition}
\noindent
\label{sec:SWVPA-def}
Let $\Delta$ be a countable alphabet
%finite (large) or infinite,
partitioned into three
subsets~$\Deltai$, $\Deltac$, $\Deltar$,
whose elements are respectively called
\emph{internal}, \emph{call} and \emph{return} symbols~\cite{AlurMadhusudan09nested}.
% names are  coined by application to functional program verification
% \begin{itemize}
% \item a set $\Deltai$ of \emph{internal symbols} denoted $a$,
% \item a set $\Deltac$ of \emph{call symbols} denoted $\call{a}$,
% \item a set $\Deltar$ of \emph{return symbols} denoted $\return{a}$.
% \end{itemize}
Let~$\< \Semiring, \oplus, \zero, \otimes, \one>$ be a commutative and complete semiring and let
$\bar\Phi = \< \Phii, \Phic, \Phir, \Phici, \Phicc, \Phicr>$
be a label theory over $\Semiring$
%In order to simplify notations, %and following the definition of Section~\ref{section:transducer},
%we shall write respectively
where $\Phii$, $\Phic$, $\Phir$ and~$\Phicx$ (with $\mathsf{x} \in \{ \mathsf{i}, \mathsf{c}, \mathsf{r}\}$)
stand respectively
for~$\Phi_\Deltai$, $\Phi_\Deltac$, $\Phi_\Deltar$ and~$\Phi_{\Deltac, \Delta_\mathsf{x}}$.
%
%Moreover, we extend this theory with a set $\Phii$
%of unary functions in $\Deltai \to \Semiring$,
%closed under $\oplus$ and $\otimes$.
